\documentclass[modern, letterpaper]{aastex61}

% TODO:
% -

% style notes
% -----------
% - Use the Makefile; don't be typing ``pdflatex'' or some bullshit.
% - Line break between sentences to make the git diffs readable.
% - Use \, as a multiply operator.
% - Reserve () for function arguments; use [] or {} for outer shit.
% - Use \sectionname not Section, \figname not Figure, \documentname not Article

\include{gitstuff}
\include{preamble}

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}1}
\newcommand{\tgas}{\acronym{TGAS}}

\begin{document}

\title{Co-moving stars in \textsl{Gaia DR1}:
       a catalog of genuine comoving star pairs from low-resolution
       spectroscopic follow-up}

\author{Adrian~M.~Price-Whelan}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}
\affiliation{To whom correspondence should be addressed:
             \texttt{adrn@astro.princeton.edu}}

\author{Semyeong~Oh}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

\author{David~N.~Spergel}
\affiliation{Flatiron Institute,
             Simons Foundation,
             162 Fifth Avenue,
             New York, NY 10010, USA}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

% \author{Melissa~Ness}
% \affiliation{todo}

% \author{David~W.~Hogg}
% \affiliation{todo}

\begin{abstract}
% Context
Widely-separated, co-eval stars are [...] for studying the chemical homogeneity
of birth, calibrating stellar atmosphere models across a range of [..], and for
modeling the Galactic tidal field.
Previous work using Hipparcos + literature RVs -> a few hundred genuine pairs
out to 8 pc.
% Aims
Identified candidate pairs using the Tycho-Gaia Astrometric Solution catalog,
Gaia DR1.
Want to identify genuinely dynamically associated pairs for future
high-resolution followup
% Methods
Low-resolution spectrograph to measure relative radial velocities to identify
genuine pairs.
% Results
Of the XX pairs observed, we find YY ...
% Conclusions
Gaia DR2 -> how many do we expect?
\end{abstract}

\keywords{
    TODO
}

\section{Introduction}\label{sec:introduction}

Outline:
- Stars form in clusters, chemically homogeneous at some level
- Clusters generally disrupt slowly through a process that preserves initial
  clustering in conserved quantities but destroys phase coherence
- Most significant process is probably tidal heating
- Dissolved clusters are therefore useful dynamical tracers of mass distribution
- In Galactic halo, globular clusters
- In disk, hope is to use chemical signatures to connect stars, then can use
  dynamics to reconstruct birth clusters, infer mass distribution / evolution
- Clusters have advantage of many stars per cluster, but fewer clusters
- Wide binary stars are little potentiometers themselves - only N=2, but are far
  more numerous as a population
- Long-term goal is to study the separation distribution of wide-binaries as a
  function of age and Galactic orbit distribution (disk vs. halo)
    - Is there a separation truncation?
    - Can that be described just from Galactic tides, or need perturbers?
- Also useful for stellar models: should have same age, chemistry, but need to
  test this
    - Chemical homogeneity tested with few open clusters - wide binaries another
      test
    - Need sample of confirmed widely-separated binaries, follow-up to get
      detailed abundances
    - Need way to verify that widely separated stars are physically associated,
      via kinematics
- Some bullshit about Gaia, TGAS, etc., released as a part of \gaia\ \DR{1}\

Most stars form in clusters with small dispersions in position, velocity, and
(likely) small initial spreads in chemical abundances.
Once a given cluster is visibly dispersing or fully dispersed, the former member
stars retain chemical and kinematic signatures of their birth cluster

comoving not wide binary (don't know if they are bound)

, though
the kinematic mapping is non-trivial because of phase-mixing.

become
useful

Widely-separated comoving stars are expected to form from the dissolution of
birth star clusters


\section{Data}\label{sec:data}

\subsection{Sample selection}\label{sec:sample}

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{tgas}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
For a full description of the selection method, see \citealt{Oh:2017}.
Briefly, likelihood 1, $L_1$, considers a model in which the full-space
velocities of each star in the pair are identical and sampled from a disk-like,
Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
in which the full-space velocities of each star were drawn independently from
the same velocity distribution used above; the likelihood ratio is computed as
$L_1/L_2$.
After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
compute the likelihood ratio for all stars within $10~\pc$ of each target star
from the \tgas\ catalog.
The likelihoods appropriately take into account the reported covariances and
uncertainties between the astrometric measurements in the \tgas\ catalog and
naturally handle geometric or projection effects in comparing spherical velocity
components for star pairs separated by large angles on the sky.
After applying a conservative cut on the likelihood ratio motivated by computing
the same likelihood ratio for random pairings of stars, we find 13,058
comoving star pairs with physical separations between $10^3~\au$ and $10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
paired with other stars.

To select candidate pairs for follow-up spectroscopic observations, we only
consider mutually exclusive star pairs within $200~\pc$ in Heliocentric
distance; with the SNR cut used above and given the limitations of the \tgas\
catalog itself, \tgas\ is $\approx$90\% complete to $200~\pc$
(\citealt{Bovy:2017}).
From this source target list, we randomly observed targets with airmass $\sec z
< 1.5$.

\subsection{Observations and Data reduction}\label{sec:reduction}

Spectra were obtained using the \project{Modspec} spectrograph mounted on the
$2.4~{\rm m}$ Hiltner telescope at MDM
observatory\footnote{\url{http://mdm.kpno.noao.edu}} on Kitt Peak (Arizona).
\project{Modspec} was set up in long-slit mode with a 1'' slit, a $600~{\rm
line}~{\rm mm}^{-1}$ grating, and a $2048\times2048~{\rm pixel}$ SITe CCD
detector (``Echelle'').
Only the central 300 pixel columns were read out from the CCD, and pixel rows
$>1600$ (at the blue end) were later trimmed because of the poor quantum
efficiency of the CCD at wavelengths $\lesssim 3600~{\rm \AA}$.
The resolution and wavelength range were $\approx2~{\rm \AA}~{\rm pixel}^{-1}$
from $\approx 3600$--$7200~{\rm \AA}$.
At this resolution, to obtain high signal-to-noise spectra, most exposures were
between $30$--$120~{\rm s}$ (15th and 85th percentiles).
To maximize the number of observed pairs, we therefore chose to not take
comparison lamp spectra at each pointing, which adds an overhead of $\approx
90~{\rm s}$ per observation.
To determine rough nightly wavelength solutions, spectra were obtained using
Hg-Ne and Ne comparison lamps at the beginning and end of each night, and
atmospheric emission lines were used to correct the wavelength solution on a
per-object basis.

We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
2017-03-15, corresponding to 321 unique star pairs and $\approx 120$ calibration
targets.
The calibration targets have previously measured radial velocities and chemical
abundances, which will be used for validation of our reduction pipeline and in
future work to measure metallicities and stellar parameters for each observed
source using \project{The Cannon} (\citealt{Ness:TODO}).
The spectra were reduced and calibrated using a custom, publicly-available
pipeline written in \python,\footnote{Released along with the code used to
generate the figures in this document at
\url{https://github.com/adrn/GaiaPairsFollowup}.} summarized below.

For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, the 2D spectra are extracted to 1D using an
empirically-estimated line spread function (LSF): at each of 16 pixel rows
between indices 750 and 850 (i.e. around the effective center of the dispersion
axis), a model for the LSF is fit to the 1D trace.
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept at
the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from each
pixel row using \texttt{L-BFGS-B} optimization (\citealt{Zhu:1994}), implemented
in the \package{scipy} package.
From the 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
At each of 1600 rows in each source image, the LSF amplitude and centroid, and
background slope and intercept at line centroid are then fit using the same
procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

\package{celerite} (\citealt{Foreman-Mackey:2017})

See: http://iopscience.iop.org/article/10.1088/0004-6256/146/6/161/pdf for an example

TODO:
- Describe pipeline...

- Comparing repeat visits on subsequent nights, systematic ~20 km/s
    - see notebook "Compare RVs with previous measurements"

\section{Results}

TODO:
- Plot of magnitude of 3d velocity difference vs. 3d separation (inferred).

\subsection{Catalog ...}\label{sec:}

\subsection{Physical properties of genuine comoving pairs}\label{sec:}

\subsection{Galactic orbits of the co-moving pairs}\label{sec:}

- Orbits of true pairs in the Galaxy, any halo or thick-disk?

\section{Conclusions }

\acknowledgements
John Brewer (Yale),
Jason Curtis (Columbia),
James Davenport (WWU) for PyDIS,
David W. Hogg (NYU),
Melissa Ness (MPIA)

This work is based on observations obtained at the MDM Observatory, operated by
Dartmouth College, Columbia University, Ohio State University, Ohio University,
and the University of Michigan.
% The authors are pleased to acknowledge that the work reported on in this paper
% was substantially performed at the TIGRESS high performance computer center at
% Princeton University which is jointly supported by the Princeton Institute for
% Computational Science and Engineering and the Princeton University Office of
% Information Technology's Research Computing department.

\software{
The code used in this project is available from
\url{https://github.com/adrn/GaiaPairsFollowup} under the MIT open-source
software license.
This research utilized the following open-source \python\ packages:
    \package{Astropy} (\citealt{Astropy-Collaboration:2013}),
    \package{astroquery} (\citealt{Ginsburg:2016}),
    \package{ccdproc} (\citealt{Craig:2015}),
    \package{celerite} (\citealt{Foreman-Mackey:2017}),
    \package{IPython} (\citealt{Perez:2007}),
    \package{matplotlib} (\citealt{Hunter:2007}),
    \package{numpy} (\citealt{Van-der-Walt:2011}),
    \package{scipy} (\url{https://www.scipy.org/}),
    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/})
}

\facility{MDM: Hiltner (Modspec)}

\bibliographystyle{aasjournal}
\bibliography{refs}

\end{document}
